---
title: "Workshop 6"
author: "YH"
date: "`r Sys.Date()`"
output: 
  github_document:
    toc: true
    toc_depth: 2
    fig_width: 10
    fig_height: 4
    df_print: kable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## Location of bus stops in Singapore

In the first workshop, we will create a data map of the location of bus stops in Singapore. We will need two pieces of information:

+ The shape file for geographical boundaries are available from [data.gov.sg](../data/BusStopLocation.zip). Download and store it in the data directory.

+ The shape file for bus stop locations is available as one of LTA's static data sets. We can download and unzip the file with `R`.

```{r}
# Download the shapefile.
download.file("https://datamall.lta.gov.sg/content/dam/datamall/datasets/Geospatial/BusStopLocation.zip",
              destfile="../data/BusStopLocation.zip")
```

Then we unzip the file by clicking the object we downloaded.

Now we are ready to recreate the map on bus stops.

```{r}
library(sf)
library(tidyverse)
busstop <- st_read("../data/BusStopLocation/BusStopLocation_Jul2023/BusStop.shp")
planningarea <- st_read("../data/MasterPlan2019RegionBoundaryNoSeaGEOJSON.geojson")
```

```{r}
ggplot(data = planningarea) +
  geom_sf(aes(geometry = geometry), fill = "steelblue", alpha = 0.3, color = "white") +
  geom_sf(data = busstop, aes(geometry = geometry), size = 1, color = "steelblue") +
  theme_minimal()
```

## Real-time Carpark Availability

Let's visualize the carpark availability data from LTA. 

```{r}
library(httr); library(jsonlite)
# Construct the resource URL
resource_url <- "http://datamall2.mytransport.sg/ltaodataservice/CarParkAvailabilityv2"
# Make the GET() request
res <- GET(resource_url, 
           add_headers(AccountKey = Sys.getenv("LTA_KEY"),
                       accept = "application/json"))
# Parse the returned data
res_content <- content(res, as = "text")
res_list <- fromJSON(res_content, flatten = TRUE)
# Extract information from sub-list
df_carpark <- as_tibble(res_list$value)
head(df_carpark)
```

```{r}
library(leaflet)
# Extract information from sublist
df_carpark <- as_tibble(res_list$value) %>%
  separate(Location, into = c("lat", "long"), sep = " ", convert = TRUE)

leaflet(df_carpark) %>%
  addTiles() %>%
  addCircleMarkers(lng = ~long, lat = ~lat,
             popup = ~paste0("Available slots: ", AvailableLots),
             radius = ~AvailableLots/100, stroke = FALSE, fillOpacity = 0.5)
```







